---
title: "Model Revision"
author: "Jen Baron"
date: 'May 22, 2020'
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
---

This is the same code as the original analysis, but I'm now modelling the raw amount of buckthorn (instead of the relative abundance)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(plyr)
library(tidyverse)
library(nlme)
library(lme4)
library(glmmTMB)
library(RColorBrewer)
library(ggrepel)
library(gridExtra)
library(emmeans)
library(predictmeans)
library(dplyr)
```


# Create Ecological Integrity Index

## Read & Join EI Data
```{r}
gap.area <- read.csv("data/gap_area.csv") %>% select(-X, -Gap)
names(gap.area)[4] <-"Gap Area"

dist.edge <- read.csv("data/dist_edge.csv")
names(dist.edge)[4]<-"Edge Distance"

frag.size <-read.csv("data/frag_size.csv")
names(frag.size)[2] <- "Fragment Size"

shade.tol <- read.csv("data/shade_tol.csv") %>% select(-X, -Gap, -tot.BA)
names(shade.tol)[4] <- "Shade Tolerance"

tree <- read.csv("data/percent_t.csv") %>% select(-X) 
names(tree)[4] <- "Percent Tree"

mgmt <- read.csv("data/management.csv")
```

```{r}
EI <- full_join(dist.edge, gap.area, by = c("Location", "Category", "ID"))
EI2 <- full_join(EI, frag.size, by = c("Location"))
EI3 <- full_join(EI2, shade.tol, by = c("Location", "Category", "ID"))
EI4 <- full_join(EI3, tree, by = c("Location", "Category", "ID"))
EI5 <- full_join(EI4, mgmt, by = c("Location"))
EI5$Category <- as.factor(EI5$Category)
names(EI5) <- sub(" ", ".", names(EI5)) #Replace spaces with .
```

## Create Matrix for EI Analysis

Variables:

- Edge Distance - plot specific
- Fragment Size - site specific
- Shade Tolerance - site specific
- Percent Tree - plot specific
- Management - site specific


```{r}
shade.tol.site <- EI5 %>% #summarize shade tolerance by site
  filter(Category == "Non-Gap") %>%
  group_by(Location) %>%
  summarize(Shade.Tolerance.Site = mean(Shade.Tolerance)) 


EI6 <- full_join(EI5, shade.tol.site, by = c("Location")) %>% select(-Shade.Tolerance)

gap.area.site <- EI5 %>% #summarize gap area
  filter(Category != "Non-Gap") %>%
  group_by(Location) %>% 
  summarize(Gap.Area.Site = mean(Gap.Area)) 

EI7 <- full_join(EI6, gap.area.site, by = c("Location")) %>% select(-Gap.Area)

head(EI7)
```

```{r}
#write.csv(EI7, file = "EI.csv")
```

## PCA

### Run PCA

Run without and without gap area

```{r}
EI7 %>% select(Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site) -> EI.data

prcomp(EI.data, scale. = TRUE) -> EI.pca

summary(EI.pca)

biplot(EI.pca)

PC1 <- predict(EI.pca)[,1]
PC2 <- predict(EI.pca)[,2]

EI7 %>% select(Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site, Gap.Area.Site) %>% 
  prcomp(scale. = TRUE) -> EI.pca2

summary(EI.pca2)

biplot(EI.pca2)

PC1b <- predict(EI.pca2)[,1]
PC2b <- predict(EI.pca2)[,2]
```


```{r}
rescale.U <- rep(EI.pca$sdev, each = length(EI.pca$sdev)) #get lengths

U.scale2 <- EI.pca$rotation * rescale.U #multiply lengths by sqrt SD

round(U.scale2^2,2) #variability in each variable for each PC
```


### QQ plots for PCA
```{r}
qqnorm(EI.data[,1]);qqline(EI.data[,1])
qqnorm(EI.data[,2]);qqline(EI.data[,2])
qqnorm(EI.data[,3]);qqline(EI.data[,3])
qqnorm(EI.data[,4]);qqline(EI.data[,4])
```




### PCA BiPlot

```{r}
#display.brewer.all(colorblindFriendly = TRUE)
mypalette <-brewer.pal(11,"BrBG")[c(1,3,7,5,9,11)]
#mypalette <- c("","","lightseagreen", "lightskyblue2", "dodgerblue",  "dodgerblue4")
#image(1:9,1,as.matrix(1:9),col=mypalette,ylab="",xaxt="n",yaxt="n",bty="n")
```

```{r}
U <- data.frame(EI.pca$rotation)
colnames(U) <- colnames(EI.pca$rotation)
rownames(U) <- rownames(EI.pca$rotation)
U$descriptors <- rownames(U)
F.1 <- data.frame(EI.pca$x) # The book calls this matrix F but I use F.1 because in R, F is shorthand for FALSE
colnames(F.1) <- colnames(EI.pca$x)
rownames(F.1) <- rownames(EI.pca$x)
str(U)
levels(U$descriptors) <- c("Distance from Forest Edge", "Forest Fragment Size", "Tree Regeneration", "Successional Stage")
U$descriptors <- as.factor(U$descriptors)
str(U)
F.1$Location<- EI7$Location 
F.1$Category <- EI7$Category 


F.1$Location = factor(F.1$Location, levels=c("Field Research Station", "Code Farm", "Westminster Ponds", "Five Points Forest", "Medway Valley", "Meadowlily Woods"))
F.1$Category = factor(F.1$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))

#Change Names
F.1$Location <- revalue(F.1$Location, c("Field Research Station"="Private 1"))
F.1$Location <- revalue(F.1$Location, c("Code Farm"="Private 2"))
F.1$Location <- revalue(F.1$Location, c("Westminster Ponds"="Public 1"))
F.1$Location <- revalue(F.1$Location, c("Five Points Forest"="Private 3"))
F.1$Location <- revalue(F.1$Location, c("Medway Valley"="Public 2"))
F.1$Location <- revalue(F.1$Location, c("Meadowlily Woods"="Public 3"))
levels((F.1$Location))

```

```{r}
levels(U$descriptors) <- c("Distance from Edge", "Fragment Size", "Tree Regeneration", 
                           "Successional Stage")


biplot1 <- ggplot(F.1, aes(x = PC1, y =PC2)) + 
  geom_point(aes(shape = Category, fill = Location),col = "black", size = 2, pch=21, alpha = 0.9) +
  theme_classic() +
  coord_fixed() +
  labs(x = 'Principle Component 1', y = "Principle Component 2") +
  scale_fill_manual(values = mypalette) +
  geom_segment(data = U, aes(xend = PC1*3, yend = PC2*3,x = 0, y = 0), col = "black", alpha = 0.7, arrow = arrow(length = unit(0.35, "cm"))) +
  geom_label_repel(data = U, aes(x = PC1*3, y = PC2*3, label = descriptors),
                   col = "black", nudge_y = -0.35, 
                   segment.colour = NA, size = 3, alpha = 0.7) +
  theme(legend.position="bottom", legend.box = "horizontal", plot.margin=grid::unit(c(0,0,0,0), "mm"))
biplot1


ggsave('figures/fig.biplotBrBu.jpeg',biplot1, units = 'cm', width = 14, height =12, dpi=500)
```

# Preliminary Modelling

## Gap Size Model

```{r}
names(gap.area) <- sub(" ", ".", names(gap.area)) #Replace spaces with .
gap.test <- lme(Gap.Area~Category,random=~1|Location,data=gap.area)

anova(gap.test)
summary(gap.test)

gap.area %>% group_by(Category) %>% summarize(gap.area = mean(Gap.Area))

intervals(gap.test,  which = "fixed")
```

```{r}
ggplot(gap.area, aes(x = Category, y = Gap.Area)) +
  geom_boxplot() +
    geom_jitter(alpha=0.6, aes(col = Location)) +
  theme_classic()
```

### Test Assumptions

Here are the residuals and fitted values:

```{r}
gap.area$mfit <- fitted(gap.test, level = 0) 

gap.area$mresid <- residuals(gap.test, type = "normalized")
```


**Linearity and Equal Variance**
```{r}
ggplot(gap.area, aes(x = mfit, y = mresid)) + 
  geom_point() +
  geom_smooth()
```

**Normality**
```{r}
ggplot(gap.area, aes(sample = mresid)) + geom_qq() + geom_qq_line()
```


**Cook's Distances**
```{r}
gap.area$cookd<-CookD(gap.test, plot=FALSE)

ggplot(gap.area, aes(seq_along(cookd), cookd))+
  geom_bar(stat="identity", position="identity")+ 
  xlab("Obs. Number")+
  ylab("Cook's distance")+
  theme_classic()
```


## Prepare Data for Buckthorn Model

```{r}
shrubs.f <- read.csv("data/shrubs_final.csv") %>% select(-"X")
shrubs.EI <- full_join(shrubs.f, EI7, by = c("Location", "Category", "ID")) 
shrubs.EI$Category <- as.factor(shrubs.EI$Category)
names(shrubs.EI) <- sub(" ", ".", names(shrubs.EI)) #Replace spaces with .
str(shrubs.EI)
```


Add values from PCA
```{r}
shrubs.EI$PC1 <- PC1
shrubs.EI$PC2 <- PC2
```

Standardize EI
```{r}
hist(shrubs.EI$PC1)
shrubs.EI %>% summarize(mean = mean(PC1) %>% round(1), sd = sd(PC1) %>% round(1)) #mean already was 0 (centered)
shrubs.EI$PC1.s <- scale(shrubs.EI$PC1, center = TRUE, scale = TRUE)  #standardize
shrubs.EI %>% summarize(mean = mean(PC1.s) %>% round(1), sd = sd(PC1.s)) 
head(shrubs.EI)
```

### Compare Private & Public

```{r}
lm.mgmt <- lm(data = shrubs.EI, PC1 ~ Management)
summary(lm.mgmt)
anova(lm.mgmt)
confint(lm.mgmt)

ggplot(shrubs.EI, aes(x = Management, y = PC1)) +
  geom_boxplot() +
    geom_jitter(alpha=0.6, aes(col = Location)) +
  theme_classic()
```



### Summarize EI

```{r}
EI.location <- shrubs.EI %>% group_by(Location) %>%
  summarize(EI = mean(PC1) %>% round(2))
arrange(EI.location, EI)

EI.cat <- shrubs.EI %>% group_by(Category) %>%
  summarize(EI = mean(PC1) %>% round(2))
arrange(EI.cat, EI)
```

### Check Buckthorn Distribution

```{r}
hist(shrubs.f$Buckthorn) #left skewed
(sum(shrubs.EI$Buckthorn > 0) / nrow(shrubs.EI))*100 # 56% zeros
```


### Set Reference Condition

Is it more interesting to consider EAB gaps or non-gaps as the reference category? Decided on EAB gaps


```{r}
#shrubs.EI$Category <- relevel(shrubs.EI$Category, ref = "Non-Gap") #re-level to set non-gap as reference condition - removed
```

### Table 1

Site level summary statistics for each gap category (EAB gap, other gap, non-gap)

```{r}
str(shrubs.EI)
shrubs.summary <- shrubs.EI %>% select(Location, Category, ID, Buckthorn, Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site, PC1) %>% 
  group_by(Location, Category) %>% 
  summarize(Buckthorn = mean(Buckthorn), Edge.Distance = mean(Edge.Distance), Fragment.Size = mean(Fragment.Size), 
            Percent.Tree = mean(Percent.Tree), Shade.Tolerance.Site = mean(Shade.Tolerance.Site), PC1 = mean(PC1)) %>% 
  mutate_if(is.numeric, round, digits = 1)
shrubs.summary

shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Field Research Station"="Private 1"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Code Farm"="Private 2"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Westminster Ponds"="Public 1"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Five Points Forest"="Private 3"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Medway Valley"="Public 2"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Meadowlily Woods"="Public 3"))

#write.csv(shrubs.summary, file = "shrubs_summary.csv")

```

```{r}
gap.area.summary <- EI5 %>% #summarize gap area
  filter(Category != "Non-Gap") %>%
  group_by(Location, Category) %>% 
  summarize(Gap.Area.Site = mean(Gap.Area) %>% round(1)) 

gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Field Research Station"="Private 1"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Code Farm"="Private 2"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Westminster Ponds"="Public 1"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Five Points Forest"="Private 3"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Medway Valley"="Public 2"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Meadowlily Woods"="Public 3"))

#write.csv(gap.area.summary, file = "gaps_summary.csv")
```

### Summarize Zeros

```{r}
shrubs.EI %>% filter(percentB == 0) %>%
  group_by(Category) %>% 
  summarise(count = n())

shrubs.EI %>% filter(percentB > 0) %>%
  group_by(Category) %>% 
  summarise(count = n())
```


## Apply a Poission Generalized MM

Check for overdispersion

```{r}
shrubs.B1 <- glmer(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = poisson, data = shrubs.EI)

E1 <- resid(shrubs.B1, type = "pearson")
N <- nrow(shrubs.EI)
p <- length(fixef(shrubs.B1)) + 1
sum(E1^2) / (N - p)
```

Yea, that's way over 1 (super overdispersed)


## Apply a Negative Binomial GLMM

Check for overdispersion
```{r}
shrubs.B2 <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)

E3 <- resid(shrubs.B2, type = "pearson")
N <- nrow(shrubs.EI)
p <- length(fixef(shrubs.B2)$cond) + 1 + 1
sum(E3^2) / (N - p)
```

Now likely underdispersed, but better than before


### Model Selection

Select ecological integrity (PC1 / PC2) with and without gap area

```{r}
#PC1 without gap area
shrubsfit.Ba <- glmmTMB(Buckthorn ~ Category * PC1 + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC1 with gap area
shrubsfit.Bb <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC2 without gap area
shrubsfit.Bc <- glmmTMB(Buckthorn ~ Category * PC2 + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC2 with gap area
shrubsfit.Bd <- glmmTMB(Buckthorn ~ Category * PC2 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
#gap area in PCA
shrubsfit.Be <- glmmTMB(Buckthorn ~ Category * PC1b + (1 | Location), family = "nbinom2", data = shrubs.EI)
```

```{r}
AIC(shrubsfit.Ba, shrubsfit.Bb, shrubsfit.Bc, shrubsfit.Bd, shrubsfit.Be)
```

PC1 with gap area (shrubsfit.Bb) is best


```{r}
shrubsfit.Bb <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
shrubsfit.Bb.no.inter <- glmmTMB(Buckthorn ~ Category + PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
shrubsfit.null <- glmmTMB(Buckthorn ~ 1 + (1 | Location), family = "nbinom2", data = shrubs.EI)
anova(shrubsfit.Bb, shrubsfit.Bb.no.inter)
AIC(shrubsfit.Bb, shrubsfit.Bb.no.inter, shrubsfit.null)
```

p-value for interaction is now 0.06 (still marginally significant)


```{r}
summary(shrubs.B2)
```

Effect of gap area now significant


### Model Validation

**Residuals vs fitted values X**

There are two types of fitted values for a GLM - the "link" predictions are estimates of $\eta$ and the response predictions are estimates of $\mu$.

```{r}
shrubs.EI$res <- residuals(shrubsfit.Bb, type = "pearson")
shrubs.EI$fit <- predict(shrubsfit.Bb, type = "response")

ggplot(shrubs.EI,aes(x = fit, y = res)) + 
  geom_point() + geom_smooth()

ggplot(shrubs.EI, aes(x = PC1, y = res)) + 
  geom_point() + geom_smooth()
```



# Zero-Inflated Model 

Terms were considered important and remained in the model if removal caused an increase in AIC of 2 or more (ΔAIC ≥ 2). 

All AIC calculations below are based on: 

$$ ΔAIC  = AIC_{Remove} - AIC_{Include}$$ 

In some cases, removing the term decreases AIC (ΔAIC < 0)


```{r}
shrubs.0infl <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location) + Management, ziformula = ~., family = "nbinom2", data = shrubs.EI)

shrubs.hurdle <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location) + Management, ziformula = ~., family = "truncated_nbinom2", data = shrubs.EI)

#summary(shrubs.0infl)
#summary(shrubs.hurdle)
```

## Model Selection


Buckthorn occurrence
- m1a: Category*PC1
- m1b: Management
- m1c: Gap.Area.Site

```{r}
full.model <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)


m1a <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1  + Management + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m1b <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category * PC1  +  Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m1c <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category * PC1  + Management + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

drop1(full.model)
AIC(m1a, m1b, m1c)
```




Buckthorn occurrence:

- m2a: Gap category
- m2b: PC1
- m2c: Management
- m2d: Gap.Area.site


```{r}
m2a <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~  PC1  + Management + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m2b <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category +  Management + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m2c <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1   + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m2d <- glmmTMB(Buckthorn ~ Category * PC1  + Management + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1  + Management + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

drop1(m1a)

AIC(m2a, m2b, m2c, m2d)
```


Buckthorn occurrence:

- m3a: Gap category
- m3b: PC1
- m3c: Management


```{r}

m3a <- glmmTMB(Buckthorn ~ Category * PC1  + Management +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~  PC1  + Management  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m3b <- glmmTMB(Buckthorn ~ Category * PC1  + Management +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~  Category  + Management  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m3c <- glmmTMB(Buckthorn ~ Category * PC1  + Management +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~  Category + PC1 + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
drop1(m2d)
AIC(m3a, m3b, m3c)
```

```{r}

m3d <- glmmTMB(Buckthorn ~ Category * PC1   + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1  + Management + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m4a <- glmmTMB(Buckthorn ~ Category * PC1  +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~  PC1  + Management  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m4b <- glmmTMB(Buckthorn ~ Category * PC1  +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category   + Management  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m4c <- glmmTMB(Buckthorn ~ Category * PC1  +  Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1   + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
drop1(m3d)
AIC(m4a, m4b, m4c)
```

```{r}
m4d <- glmmTMB(Buckthorn ~ Category * PC1   + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category + PC1 + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m5a <- glmmTMB(Buckthorn ~ Category * PC1   + Gap.Area.Site + (1 | Location), 
                    ziformula = ~  PC1 + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
m5b <- glmmTMB(Buckthorn ~ Category * PC1   + Gap.Area.Site + (1 | Location), 
                    ziformula = ~ Category  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

drop1(m4d)
AIC(m5a, m5b)
```

```{r}
m6a <- glmmTMB(Buckthorn ~ Category * PC1   +  (1 | Location), 
                    ziformula = ~ 1  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

drop1(m5b)
AIC(m6a)
```

```{r}
m6b <- glmmTMB(Buckthorn ~ Category * PC1  + (1 | Location), 
                    ziformula = ~  Category + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

m7a <- glmmTMB(Buckthorn ~ Category * PC1  + (1 | Location), 
                    ziformula = ~  1   + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
drop1(m6b)
AIC(m7a)
```

```{r}
shrubs.hurdle5 <- glmmTMB(Buckthorn ~ Category * PC1  + (1 | Location), 
                    ziformula = ~  Category + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
```



## Plot Model Predictions


```{r}
non.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "Non-Gap"))
non.max<- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "Non-Gap"))
EAB.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "EAB Gap"))
EAB.max <- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "EAB Gap"))
other.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "Other Gap"))
other.max <- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "Other Gap"))


pred <- data.frame(PC1 = c(seq(EAB.min, EAB.max, length.out = 100), 
                               seq(other.min, other.max, length.out = 100), 
                                   seq(non.min, non.max, length.out = 100)),
                   Category = c(rep("EAB Gap", 100), rep("Other Gap", 100), rep("Non-Gap", 100)),
                   Gap.Area.Site = rep(mean(shrubs.EI$Gap.Area.Site),300), # Added gap area
                   Management = c(rep("Private", 150), rep("Public", 150)), 
                   Location = rep(NA, 300)) # This NA gives general predictions instead of location specific predictions

pred$fit <- predict(shrubs.hurdle5, newdata = pred, se.fit = TRUE, type = "link")$fit
pred$se <- predict(shrubs.hurdle5, newdata = pred, se.fit = TRUE, type = "link")$se.fit
str(pred)

pred.upr <- pred$fit + 1.96 * pred$se
pred.lwr <- pred$fit - 1.96 * pred$se

pred$upr <- ifelse(exp(pred.upr) < 100, exp(pred.upr), 100)
pred$lwr <- ifelse(exp(pred.lwr) > 0, exp(pred.lwr), 0)

pred$mean <- exp(pred$fit)

```


```{r}
mypalette2 = c(brewer.pal(12, "Paired")[c(12,2,4)])
```

Figure for publication (no legend)

```{r}
shrubs.EI$Category = factor(shrubs.EI$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))
pred$Category = factor(pred$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))


fig.EI <- ggplot(pred, aes(x = PC1, y = mean)) +
    geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
      geom_point(data = shrubs.EI, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity (index)",
       y = "European buckthorn (% cover)",
       col = "Category") +
  scale_y_continuous(limits = c(0, 100)) +
    scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
    theme_classic() +
 theme(legend.position = "none")

fig.EI

ggsave('figures/fig.EI.jpeg',fig.EI , units = 'cm', width = 14, height = 10, dpi=500)

```

Graphical abstract

```{r}
fig.EI2 <- ggplot(pred, aes(x = PC1, y = mean)) +
  geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
  geom_point(data = shrubs.EI, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity (index)",
       y = "European buckthorn (% cover)",
       col = "Category") +
  scale_y_continuous(limits = c(0, 100)) +
  scale_colour_manual(values = mypalette2,  name = "Gap Category", labels = c("Emerald Ash Borer Gap", "Other Canopy Gap", "Closed Canopy")) +
  scale_fill_manual(values = mypalette2,  name = "Gap Category", labels = c("Emerald Ash Borer Gap", "Other Canopy Gap", "Closed Canopy")) +
  ggtitle("Emerald-ash borer interacts with ecological integrity to facilitate \nEuropean buckthorn through the creation of canopy gaps") +
  theme_classic() +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"), plot.margin=unit(c(1,0.5,0,.5),"cm"))

fig.EI2

#ggsave('figures/fig.EI2.jpeg',fig.EI2 , width = 22, height = 14, units = "cm")

```


## Model Validatation 

### Through Simulation

```{r}

sim.m5 <- simulate(shrubs.hurdle5, nsim = 999) %>% t() %>% as.data.frame() %>% 
  gather(observation, sim_value)

sim.m5$Location <- rep(shrubs.EI$Location, each = 999)
sim.m5$obs.num <- as.numeric(str_extract(sim.m5$observation, "[:digit:]+"))


str(sim.m5)


ggplot() +
  geom_boxplot(data = sim.m5, aes(x = reorder(observation, obs.num), 
                                  y = sim_value + 1, 
                                  col = Location), alpha = 0.5) +
  scale_y_log10() +
  geom_point(data = shrubs.EI, aes(x = 1:72, y = Buckthorn + 1), col = "black") +
  theme_classic()

```

A lot of uncertainty but the real observations are usually in line the middle 50% of the simulations.  

A bit more buckthorn than the model predicts at Westminster Ponds but I think that is consistent with previous knowledge.


### Examine High % Buckthorn Points

Check those high % buckthorn points - where are they coming from?

```{r}
ggplot(pred, aes(x = PC1, y = mean)) +
  geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
  geom_point(data = shrubs.EI, aes(x = PC1, y = Buckthorn, col = Category, shape = Location), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
  theme_classic() 
# theme(legend.position = "none")

shrubs.EI.w <- shrubs.EI %>% filter(Location == "Westminster Ponds")
shrubs.EI.nw <- shrubs.EI %>% filter(Location != "Westminster Ponds")

ggplot(pred, aes(x = PC1, y = mean)) +
  geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
  geom_point(data = shrubs.EI.nw, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_point(data = shrubs.EI.w, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6, shape = 17) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
  theme_classic() 
# theme(legend.position = "none")

```


The 90% are all coming from Westminster Ponds


### Run Model Without Western FRS

As per reviewer comments - influence of this station

Remove FRS from data
```{r}
shrubs.EI.wo <- shrubs.EI %>% filter(Location != "Field Research Station") %>% 
  select(Location, Category, ID, Buckthorn, Management, Gap.Area.Site, PC1) 
str(shrubs.EI.wo)
```

Run model without FRS data
```{r}
shrubs.hurdle5.wo <- glmmTMB(Buckthorn ~ Category * PC1 + (1 | Location), 
                        ziformula = ~ Category +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI.wo)
```

**Compare Model Outputs**
```{r}
exp(fixef(shrubs.hurdle5)$cond) # Intercept expected cover when present, others are multiplicative changes in cover conditional on presence
exp(fixef(shrubs.hurdle5.wo)$cond)
```

Remove FRS: 

- Even more buckthorn in average EI EAB gaps
- Greater discrepancy between EAB gaps & non-gaps
- Slope of PC1 in EAB gaps increases

Therefore, greater EAB & EI effect when we remove FRS 


```{r}
#Non-gaps
33.7154190 * 0.2325890
49.2315112 * 0.1975805

#Other gaps
33.7154190 * 0.5355936
49.2315112 *  0.3845417
```

```{r}
exp(fixef(shrubs.hurdle5)$zi) # Intercept is odds of zero, others are odds ratios
exp(fixef(shrubs.hurdle5.wo)$zi) # Intercept is odds of zero, others are odds ratios
```

Remove FRS:

- Still equally likely to be present in EAB & other gaps
- Odds of presense decrease from 1 to 0.76
- Odds of presence in non-gaps also decrease (5.7 to 5.0 times less likely)


```{r}
exp(confint(shrubs.hurdle5))
exp(confint(shrubs.hurdle5.wo))
```

Remove FRS:

- In general, confidence intervals become wider (more uncertainty)


```{r}
pred.wo <- data.frame(PC1 = c(seq(EAB.min, EAB.max, length.out = 100), 
                               seq(other.min, other.max, length.out = 100), 
                                   seq(non.min, non.max, length.out = 100)),
                   Category = c(rep("EAB Gap", 100), rep("Other Gap", 100), rep("Non-Gap", 100)),
                   Gap.Area.Site = rep(mean(shrubs.EI$Gap.Area.Site),300), # Added gap area
                   Management = c(rep("Private", 150), rep("Public", 150)), 
                   Location = rep(NA, 300)) # This NA gives general predictions instead of location specific predictions

pred.wo$fit <- predict(shrubs.hurdle5.wo, newdata = pred, se.fit = TRUE, type = "link")$fit
pred.wo$se <- predict(shrubs.hurdle5.wo, newdata = pred, se.fit = TRUE, type = "link")$se.fit
str(pred.wo)

pred.upr <- pred.wo$fit + 1.96 * pred$se
pred.lwr <- pred.wo$fit - 1.96 * pred$se

pred.wo$upr <- ifelse(exp(pred.upr) < 100, exp(pred.upr), 100)
pred.wo$lwr <- ifelse(exp(pred.lwr) > 0, exp(pred.lwr), 0)

pred.wo$mean <- exp(pred.wo$fit)
```



```{r}
shrubs.EI.wo$Category = factor(shrubs.EI.wo$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))
pred.wo$Category = factor(pred.wo$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))

fig.EI.wo <- ggplot(pred.wo, aes(x = PC1, y = mean)) +
    geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
      geom_point(data = shrubs.EI.wo, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_y_continuous(limits = c(0, 100)) +
    scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
    theme_classic() +
 theme(legend.position = "none")

fig.EI
fig.EI.wo

#ggsave('figures/fig.EI.noFRS.jpeg',fig.EI.wo , units = 'cm', width = 14, height = 10)
```






## Model Results

Results of zero-inflated model are odds ratios, interpreted as probability of buckthorn occurance. These results are reported relative to EAB gaps and thus are not converted to the probability of occurance. 

Results of truncated negatived binomial model are reported as expected cover when present and converted from the multiplicative changes for non-reference condition, interpreted as buckthorn abundance. The PC1*Category interaction predicts the relationship between PC1 and gap category, converted from multiplicative changes for non-reference condition.

### Model Summary

```{r}
summary(shrubs.hurdle5)
```

**Zero-Inflated (Occurance)**
```{r}
exp(fixef(shrubs.hurdle5)$zi) # Intercept is odds of zero, others are odds ratios
```


**Truncated Negative Binomial (Abundance)**
```{r}
exp(fixef(shrubs.hurdle5)$cond) # Intercept expected cover when present, others are multiplicative changes in cover conditional on presence
```

**Confidence Intervals**
```{r}

exp(confint(shrubs.hurdle5))
```

### Buckthorn Occurance

The zero-altered model predicts the odds of observing a zero. We have to convert these values from the probability of a zero to to the probability of a non-zero (subtracting from 1).

The odds of observing a zero in an EAB gap are 0.41 (0.11, 1.5)

```{r}
((1-0.4113465)*100) %>% round(1) #estimate
((1-0.1102974)*100) %>% round(1) #upper
(abs(1-1.5340883)*100) %>% round(1) #lower
```


### Buckthorn Abundance: Category

**EAB Gaps**

33.7 (22.3, 50.9)

**Other Gaps**
```{r}
#Other-Gaps
(33.7154190*0.5355936) %>% round(2)#estimate
(33.7154190*0.3387368) %>% round(2)#lower
(33.7154190*0.8468538) %>% round(2)#upper
```

**Non-Gaps**
```{r}
#Non-Gaps
(33.715419*0.2325890) %>% round(2) #estimate
(33.715419*0.1132062) %>% round(2)#lower
(33.715419*0.4778683) %>% round(2)#upper
```

### Buckthorn Abundance: PC1*Category

**EAB Gaps**

0.71 (0.52, 0.97)

**Other Gaps**
```{r}
#Other-Gaps
(0.7128030*1.7871821) %>% round(2)#estimate
(0.7128030*1.2296463) %>% round(2)#lower 
(0.7128030*2.5975110) %>% round(2)#upper
```

**Non-Gaps**
```{r}
#Non-Gaps
(0.7128030*1.2055000) %>% round(2)#estimate
(0.7128030*0.6601346) %>% round(2)#lower
(0.7128030*2.2014151) %>% round(2)#upper
```







# Reproducibility
```{r}
Sys.time()
git2r::repository()
sessionInfo()
```

