---
title: "Model Revision"
author: "Jen Baron"
date: 'May 22, 2020'
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    keep_md: true
---

This is the same code as the original analysis, but I'm now modelling the raw amount of buckthorn (instead of the relative abundance)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(plyr)
library(tidyverse)
library(nlme)
library(lme4)
library(glmmTMB)
library(RColorBrewer)
library(ggrepel)
library(gridExtra)
library(DHARMa)
library(emmeans)
library(dplyr)
```


# Create Scores

## Read in & Join EI Data
```{r}
gap.area <- read.csv("data/gap_area.csv") %>% select(-X, -Gap)
names(gap.area)[4] <-"Gap Area"
str(gap.area)

dist.edge <- read.csv("data/dist_edge.csv")
names(dist.edge)[4]<-"Edge Distance"
str(dist.edge)

frag.size <-read.csv("data/frag_size.csv")
names(frag.size)[2] <- "Fragment Size"
str(frag.size)

shade.tol <- read.csv("data/shade_tol.csv") %>% select(-X, -Gap, -tot.BA)
names(shade.tol)[4] <- "Shade Tolerance"
str(shade.tol)

tree <- read.csv("data/percent_t.csv") %>% select(-X) #wondering if this should be absolute or relative...
names(tree)[4] <- "Percent Tree" #absolute right now (not relative to shrub cover)
str(tree)

mgmt <- read.csv("data/management.csv")
str(mgmt)
```

```{r}
EI <- full_join(dist.edge, gap.area, by = c("Location", "Category", "ID"))
EI2 <- full_join(EI, frag.size, by = c("Location"))
EI3 <- full_join(EI2, shade.tol, by = c("Location", "Category", "ID"))
EI4 <- full_join(EI3, tree, by = c("Location", "Category", "ID"))
EI5 <- full_join(EI4, mgmt, by = c("Location"))
EI5$Category <- as.factor(EI5$Category)
names(EI5) <- sub(" ", ".", names(EI5)) #Replace spaces with .
str(EI5)
```

## Create Matrix for EI Analysis

Variables:

- Edge Distance - plot specific: lower number = lower integrity
- Fragment Size - site specific: lower number = lower integrity
- Shade Tolerance - site specific: lower number = lower integrity?
- Percent Tree - plot specific: lower number = lower integrity?
- Management - site specific: lower number = lower integrity?


```{r}
shade.tol.site <- EI5 %>% #summarize shade tolerance by site
  filter(Category == "Non-Gap") %>%
  group_by(Location) %>%
  summarize(Shade.Tolerance.Site = mean(Shade.Tolerance)) 

str(shade.tol.site)

EI6 <- full_join(EI5, shade.tol.site, by = c("Location")) %>% select(-Shade.Tolerance)

gap.area.site <- EI5 %>% #summarize gap area
  filter(Category != "Non-Gap") %>%
  group_by(Location) %>% 
  summarize(Gap.Area.Site = mean(Gap.Area)) 

EI7 <- full_join(EI6, gap.area.site, by = c("Location")) %>% select(-Gap.Area)

head(EI7)
```

```{r}
#write.csv(EI7, file = "EI.csv")
```

## Run without and without gap area
```{r}
EI7 %>% select(Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site) -> EI.data

prcomp(EI.data, scale. = TRUE) -> EI.pca

summary(EI.pca)

biplot(EI.pca)

PC1 <- predict(EI.pca)[,1]
PC2 <- predict(EI.pca)[,2]

EI7 %>% select(Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site, Gap.Area.Site) %>% 
  prcomp(scale. = TRUE) -> EI.pca2

summary(EI.pca2)

biplot(EI.pca2)

PC1b <- predict(EI.pca2)[,1]
PC2b <- predict(EI.pca2)[,2]
```


```{r}
rescale.U <- rep(EI.pca$sdev, each = length(EI.pca$sdev)) #get lengths

U.scale2 <- EI.pca$rotation * rescale.U #multiply lengths by sqrt SD

round(U.scale2^2,2) #variability in each variable for each PC
```


### QQ plots for PCA
```{r}
qqnorm(EI.data[,1]);qqline(EI.data[,1])
qqnorm(EI.data[,2]);qqline(EI.data[,2])
qqnorm(EI.data[,3]);qqline(EI.data[,3])
qqnorm(EI.data[,4]);qqline(EI.data[,4])
```




### PCA BiPlot

```{r}
#display.brewer.all(colorblindFriendly = TRUE)
mypalette <-brewer.pal(11,"BrBG")[c(1,3,7,5,9,11)]
#mypalette <- c("","","lightseagreen", "lightskyblue2", "dodgerblue",  "dodgerblue4")
image(1:9,1,as.matrix(1:9),col=mypalette,ylab="",xaxt="n",yaxt="n",bty="n")
```

```{r}
U <- data.frame(EI.pca$rotation)
colnames(U) <- colnames(EI.pca$rotation)
rownames(U) <- rownames(EI.pca$rotation)
U$descriptors <- rownames(U)
F.1 <- data.frame(EI.pca$x) # The book calls this matrix F but I use F.1 because in R, F is shorthand for FALSE
colnames(F.1) <- colnames(EI.pca$x)
rownames(F.1) <- rownames(EI.pca$x)
str(U)
levels(U$descriptors) <- c("Distance from Forest Edge", "Forest Fragment Size", "Tree Regeneration", "Successional Stage")
U$descriptors <- as.factor(U$descriptors)
str(U)
F.1$Location<- EI7$Location 
F.1$Category <- EI7$Category 


F.1$Location = factor(F.1$Location, levels=c("Field Research Station", "Code Farm", "Westminster Ponds", "Five Points Forest", "Medway Valley", "Meadowlily Woods"))
F.1$Category = factor(F.1$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))

#Change Names
F.1$Location <- revalue(F.1$Location, c("Field Research Station"="Private 1"))
F.1$Location <- revalue(F.1$Location, c("Code Farm"="Private 2"))
F.1$Location <- revalue(F.1$Location, c("Westminster Ponds"="Public 1"))
F.1$Location <- revalue(F.1$Location, c("Five Points Forest"="Private 3"))
F.1$Location <- revalue(F.1$Location, c("Medway Valley"="Public 2"))
F.1$Location <- revalue(F.1$Location, c("Meadowlily Woods"="Public 3"))
levels((F.1$Location))

```

Plot the objects
```{r}
levels(U$descriptors) <- c("Distance from Edge", "Fragment Size", "Tree Regeneration", 
                           "Successional Stage")


biplot1 <- ggplot(F.1, aes(x = PC1, y =PC2)) + 
  geom_point(aes(shape = Category, fill = Location),col = "black", size = 2, pch=21, alpha = 0.9) +
  theme_classic() +
  coord_fixed() +
  labs(x = 'Principle Component 1', y = "Principle Component 2") +
  scale_fill_manual(values = mypalette) +
  geom_segment(data = U, aes(xend = PC1*3, yend = PC2*3,x = 0, y = 0), col = "black", alpha = 0.7, arrow = arrow(length = unit(0.35, "cm"))) +
  geom_label_repel(data = U, aes(x = PC1*3, y = PC2*3, label = descriptors),
                   col = "black", nudge_y = -0.35, 
                   segment.colour = NA, size = 3, alpha = 0.7) +
  theme(legend.position="bottom", legend.box = "horizontal", plot.margin=grid::unit(c(0,0,0,0), "mm"))
biplot1


#ggsave('fig.biplotBrBu.jpeg',biplot1, units = 'cm', width = 14, height =12)
```



# Modelling

## Read in Shrub Data
```{r}
shrubs.f <- read.csv("data/shrubs_final.csv") %>% select(-"X")
shrubs.EI <- full_join(shrubs.f, EI7, by = c("Location", "Category", "ID")) 
shrubs.EI$Category <- as.factor(shrubs.EI$Category)
names(shrubs.EI) <- sub(" ", ".", names(shrubs.EI)) #Replace spaces with .
str(shrubs.EI)
```


### Add values from PCA


```{r}
shrubs.EI$PC1 <- PC1
shrubs.EI$PC2 <- PC2
```

### Standardize EI

```{r}
hist(shrubs.EI$PC1)
shrubs.EI %>% summarize(mean = mean(PC1) %>% round(1), sd = sd(PC1) %>% round(1)) #mean already was 0 (centered)
shrubs.EI$PC1.s <- scale(shrubs.EI$PC1, center = TRUE, scale = TRUE)  #standardize
shrubs.EI %>% summarize(mean = mean(PC1.s) %>% round(1), sd = sd(PC1.s)) 
head(shrubs.EI)
```


### Compare Private & Public

```{r}

lm.mgmt <- lm(data = shrubs.EI, PC1 ~ Management)
summary(lm.mgmt)
anova(lm.mgmt)
confint(lm.mgmt)

ggplot(shrubs.EI, aes(x = Management, y = PC1)) +
  geom_boxplot() +
    geom_jitter(alpha=0.6, aes(col = Location)) +
  theme_classic()
```



### Summarize EI

```{r}
EI.location <- shrubs.EI %>% group_by(Location) %>%
  summarize(EI = mean(PC1) %>% round(2))
arrange(EI.location, EI)

EI.cat <- shrubs.EI %>% group_by(Category) %>%
  summarize(EI = mean(PC1) %>% round(2))
arrange(EI.cat, EI)
```

### Check distribution
```{r}
hist(shrubs.f$Buckthorn) #left skewed
(sum(shrubs.EI$Buckthorn > 0) / nrow(shrubs.EI))*100 # 56% zeros
```



```{r}
#shrubs.EI$Category <- relevel(shrubs.EI$Category, ref = "Non-Gap") #re-level to set non-gap as reference condition - removed
```

Is it more interesting to consider EAB gaps or non-gaps as the reference category? Decided on EAB gaps


## Apply a Poission Generalized MM  & check for overdispersion
```{r}
shrubs.B1 <- glmer(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = poisson, data = shrubs.EI)

E1 <- resid(shrubs.B1, type = "pearson")
N <- nrow(shrubs.EI)
p <- length(fixef(shrubs.B1)) + 1
sum(E1^2) / (N - p)
```

Yea, that's way over 1 (super overdispersed)


## Apply a negative binomial GLMM  & check for overdispersion
```{r}
shrubs.B2 <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)

E3 <- resid(shrubs.B2, type = "pearson")
N <- nrow(shrubs.EI)
p <- length(fixef(shrubs.B2)$cond) + 1 + 1
sum(E3^2) / (N - p)
```

Now likely underdispersed, but better than before


### Model Selection

Select ecological integrity (PC1 / PC2) with and without gap area

```{r}
#PC1 without gap area
shrubsfit.Ba <- glmmTMB(Buckthorn ~ Category * PC1 + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC1 with gap area
shrubsfit.Bb <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC2 without gap area
shrubsfit.Bc <- glmmTMB(Buckthorn ~ Category * PC2 + (1 | Location), family = "nbinom2", data = shrubs.EI)
#PC2 with gap area
shrubsfit.Bd <- glmmTMB(Buckthorn ~ Category * PC2 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
#gap area in PCA
shrubsfit.Be <- glmmTMB(Buckthorn ~ Category * PC1b + (1 | Location), family = "nbinom2", data = shrubs.EI)
```

```{r}
AIC(shrubsfit.Ba, shrubsfit.Bb, shrubsfit.Bc, shrubsfit.Bd, shrubsfit.Be)
```

PC1 with gap area (shrubsfit.Bb) is best


```{r}
shrubsfit.Bb <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
shrubsfit.Bb.no.inter <- glmmTMB(Buckthorn ~ Category + PC1 + Gap.Area.Site + (1 | Location), family = "nbinom2", data = shrubs.EI)
shrubsfit.null <- glmmTMB(Buckthorn ~ 1 + (1 | Location), family = "nbinom2", data = shrubs.EI)
anova(shrubsfit.Bb, shrubsfit.Bb.no.inter)
AIC(shrubsfit.Bb, shrubsfit.Bb.no.inter, shrubsfit.null)
```

p-value for interaction is now 0.06 (still marginally significant)


```{r}
summary(shrubs.B2)
```

Effect of gap area now significant


### Model Validation

#### Residuals vs fitted values X

There are two types of fitted values for a GLM - the "link" predictions are estimates of $\eta$ and the response predictions are estimates of $\mu$.

```{r}
shrubs.EI$res <- residuals(shrubsfit.Bb, type = "pearson")
shrubs.EI$fit <- predict(shrubsfit.Bb, type = "response")

ggplot(shrubs.EI,aes(x = fit, y = res)) + 
  geom_point() + geom_smooth()

ggplot(shrubs.EI, aes(x = PC1, y = res)) + 
  geom_point() + geom_smooth()
```

#### DHARMa Plots

```{r}
simulationOutput1 <- simulateResiduals(fittedModel = shrubsfit.Bb, n = 1000)
plotSimulatedResiduals(simulationOutput = simulationOutput1)
par(mfrow = c(1,3))
```


### Test Gap Size

```{r}
names(gap.area) <- sub(" ", ".", names(gap.area)) #Replace spaces with .
gap.test <- lme(Gap.Area~Category,random=~1|Location,data=gap.area)
anova(gap.test)
summary(gap.test)
gap.area %>% group_by(Category) %>% summarize(gap.area = mean(Gap.Area))
intervals(gap.test,  which = "fixed")
```

```{r}
ggplot(gap.area, aes(x = Category, y = Gap.Area)) +
  geom_boxplot() +
    geom_jitter(alpha=0.6, aes(col = Location)) +
  theme_classic()
```

### Model Selection

```{r}
shrubs.0infl <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location) + Management, ziformula = ~., family = "nbinom2", data = shrubs.EI)

shrubs.hurdle <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location) + Management, ziformula = ~., family = "truncated_nbinom2", data = shrubs.EI)

#summary(shrubs.0infl)
#summary(shrubs.hurdle)
```


**Step 1: Compare to Negative Binomial**

```{r}
AIC(shrubs.0infl, shrubs.hurdle, shrubs.B2) # Better than neg binomial 
```

Both the zero-inflated and zero-altered model are better than the original negative binomial model

We're going to use zero-altered because it's safe to assume that if buckthorn was present, it was recorded

**Step 2: Fixed Effects**

```{r}
str(shrubs.EI)
shrubs.hurdle <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + Management + (1 | Location), 
                        ziformula = ~ Category * PC1 + Gap.Area.Site + Management + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)
```

Drop Management

```{r}
#AIC
427.70-428.10 #management
429.85-427.70 #gap size
```


**Step 3: Simplify Zero-Inflation**

```{r}
shrubs.hurdle1a <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~  Category + PC1 + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

shrubs.hurdle1b <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~ Category * PC1  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)


AIC(shrubs.hurdle, shrubs.hurdle1a, shrubs.hurdle1b)

# 1a - drop interaction

shrubs.hurdle2a <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~  PC1 + Category + Gap.Area.Site + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

shrubs.hurdle2b <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~  Category + Gap.Area.Site +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

shrubs.hurdle2c <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~  Category + PC1 + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

AIC(shrubs.hurdle1a, shrubs.hurdle2a, shrubs.hurdle2b, shrubs.hurdle2c)

## 2c - drop gap area

shrubs.hurdle3a <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~   PC1 +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

shrubs.hurdle3b <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~  Category  + (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)


AIC(shrubs.hurdle2c, shrubs.hurdle3a, shrubs.hurdle3b)

# 3b - drop PC1

shrubs.hurdle4 <- glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~ 1 +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

AIC(shrubs.hurdle3b, shrubs.hurdle4)
anova(shrubs.hurdle3b, shrubs.hurdle4)

# Keep Category

drop1(shrubs.hurdle3b)
```

```{r}
#AIC Scores

#Gap Category
430.9292-426.0300

#Ecological integrity
427.5389-426.0300

#Gap size
428.8045-427.5389

#EI*Category
429.9393-428.8045

```

**Step 4: Simplify Negative-Binomial**

```{r}
shrubs.hurdle5 <- glmmTMB(Buckthorn ~ Category * PC1 + (1 | Location), 
                        ziformula = ~ Category +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)

shrubs.hurdle5b <-  glmmTMB(Buckthorn ~ Category * PC1 + Gap.Area.Site + (1 | Location), 
                        ziformula = ~ Category +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)


shrubs.hurdle6 <- glmmTMB(Buckthorn ~ Category + PC1 + (1 | Location), 
                        ziformula = ~ Category +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI)


drop1(shrubs.hurdle5b)

drop1(shrubs.hurdle5)

AIC(shrubs.hurdle5, shrubs.hurdle6)
```

Gap area is marginal and can be dropped

```{r}
#AIC Scores

#PC1*Category
427.9651-423.2072

#Gap area
423.21-422.03
```


### Model Validatation (through simulation)

```{r}

sim.m5 <- simulate(shrubs.hurdle5, nsim = 999) %>% t() %>% as.data.frame() %>% 
  gather(observation, sim_value)

sim.m5$Location <- rep(shrubs.EI$Location, each = 999)
sim.m5$obs.num <- as.numeric(str_extract(sim.m5$observation, "[:digit:]+"))


str(sim.m5)


ggplot() +
  geom_boxplot(data = sim.m5, aes(x = reorder(observation, obs.num), 
                                  y = sim_value + 1, 
                                  col = Location), alpha = 0.5) +
  scale_y_log10() +
  geom_point(data = shrubs.EI, aes(x = 1:72, y = Buckthorn + 1), col = "black") +
  theme_classic()

```

A lot of uncertainty but the real observations are usually in line the middle 50% of the simulations.  

A bit more buckthorn than the model predicts at Westminster Ponds but I think that is consistent with previous knowledge.

### Model validatation through simulation

```{r}

sim.m5 <- simulate(shrubs.hurdle5, nsim = 999) %>% t() %>% as.data.frame() %>% 
  gather(observation, sim_value)

sim.m5$Location <- rep(shrubs.EI$Location, each = 999)
sim.m5$obs.num <- as.numeric(str_extract(sim.m5$observation, "[:digit:]+"))



str(sim.m5)



ggplot() +
  geom_boxplot(data = sim.m5, aes(x = reorder(observation, obs.num), 
                                  y = sim_value + 1, 
                                  col = Location), alpha = 0.5) +
  scale_y_log10() +
  geom_point(data = shrubs.EI, aes(x = 1:72, y = Buckthorn + 1), col = "black") +
  theme_classic()


```


### Plot Model Predictions


```{r}
non.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "Non-Gap"))
non.max<- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "Non-Gap"))
EAB.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "EAB Gap"))
EAB.max <- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "EAB Gap"))
other.min <- min(subset(shrubs.EI$PC1, shrubs.EI$Category == "Other Gap"))
other.max <- max(subset(shrubs.EI$PC1, shrubs.EI$Category == "Other Gap"))


pred <- data.frame(PC1 = c(seq(EAB.min, EAB.max, length.out = 100), 
                               seq(other.min, other.max, length.out = 100), 
                                   seq(non.min, non.max, length.out = 100)),
                   Category = c(rep("EAB Gap", 100), rep("Other Gap", 100), rep("Non-Gap", 100)),
                   Gap.Area.Site = rep(mean(shrubs.EI$Gap.Area.Site),300), # Added gap area
                   Management = c(rep("Private", 150), rep("Public", 150)), 
                   Location = rep(NA, 300)) # This NA gives general predictions instead of location specific predictions

pred$fit <- predict(shrubs.hurdle5, newdata = pred, se.fit = TRUE, type = "link")$fit
pred$se <- predict(shrubs.hurdle5, newdata = pred, se.fit = TRUE, type = "link")$se.fit
str(pred)

pred.upr <- pred$fit + 1.96 * pred$se
pred.lwr <- pred$fit - 1.96 * pred$se

pred$upr <- ifelse(exp(pred.upr) < 100, exp(pred.upr), 100)
pred$lwr <- ifelse(exp(pred.lwr) > 0, exp(pred.lwr), 0)

pred$mean <- exp(pred$fit)

```


```{r}
mypalette2 = c(brewer.pal(12, "Paired")[c(12,2,4)])
str(pred)
```

```{r}
shrubs.EI$Category = factor(shrubs.EI$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))
pred$Category = factor(pred$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))


fig.EI <- ggplot(pred, aes(x = PC1, y = mean)) +
    geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
      geom_point(data = shrubs.EI, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity (index)",
       y = "European buckthorn (% cover)",
       col = "Category") +
  scale_y_continuous(limits = c(0, 100)) +
    scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
    theme_classic() +
 theme(legend.position = "none")

fig.EI

#ggsave('fig.EI.jpeg',fig.EI , units = 'cm', width = 14, height = 10)

```

Check those high % buckthorn points - where are they coming from?

```{r}
ggplot(pred, aes(x = PC1, y = mean)) +
  geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
  geom_point(data = shrubs.EI, aes(x = PC1, y = Buckthorn, col = Category, shape = Location), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
  theme_classic() 
# theme(legend.position = "none")

shrubs.EI.w <- shrubs.EI %>% filter(Location == "Westminster Ponds")
shrubs.EI.nw <- shrubs.EI %>% filter(Location != "Westminster Ponds")

ggplot(pred, aes(x = PC1, y = mean)) +
  geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
  geom_point(data = shrubs.EI.nw, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_point(data = shrubs.EI.w, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6, shape = 17) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
  theme_classic() 
# theme(legend.position = "none")

```


The 90% are all coming from Westminster Ponds

## Model Summary

```{r}
summary(shrubs.hurdle5)

exp(fixef(shrubs.hurdle5)$cond) # Intercept expected cover when present, others are multiplicative changes in cover conditional on presence


exp(fixef(shrubs.hurdle5)$zi) # Intercept is odds of zero, others are odds ratios

exp(confint(shrubs.hurdle5))

```

I think I've figured out how to correctly interpret the odds ratios and intercept scores

```{r}
#Intercept
#Non-Gaps
33.715419*0.1132062 %>% round(2)#lower
33.715419*0.2325890 %>% round(2) #estimate
33.715419*0.4778683 %>% round(2)#upper
#Other-Gaps
33.7154190*0.3387368 %>% round(2)#lower
33.7154190*0.5355936 %>% round(2)#estimate
33.7154190*0.8468538 %>% round(2)#upper

#PC1
#Non-Gaps
0.7128030*0.6601346 %>% round(2)#lower
0.7128030*1.2055000 %>% round(2)#estimate
0.7128030*2.2014151 %>% round(2)#upper
#Other-Gaps
0.7128030*1.2296463 %>% round(2)#lower 
0.7128030*1.7871821 %>% round(2)#estimate
0.7128030*2.5975110 %>% round(2)#upper
```

### emtrends

emtrends extracts the slopes which are expressed relative to EAB gaps - I then use exp to "unlink" the predictions.

These values are really just PC1 * (Category:PC1)

I think the levels in here get mixed up somehow because the other-gap and non-gap categories come out mixed up - I'm going to use the values from above but I know where they're coming from onw. 

```{r}
emtrends(shrubs.hurdle5, pairwise ~ Category, var = "PC1")

exp(-0.339) %>% round(3)
exp(-0.152) %>% round(3)
exp(0.242) %>% round(3)
```





## Summary Statistics (for table)

```{r}
str(shrubs.EI)
shrubs.summary <- shrubs.EI %>% select(Location, Category, ID, Buckthorn, Edge.Distance, Fragment.Size, Percent.Tree, Shade.Tolerance.Site, PC1) %>% 
  group_by(Location, Category) %>% 
  summarize(Buckthorn = mean(Buckthorn), Edge.Distance = mean(Edge.Distance), Fragment.Size = mean(Fragment.Size), 
            Percent.Tree = mean(Percent.Tree), Shade.Tolerance.Site = mean(Shade.Tolerance.Site), PC1 = mean(PC1)) %>% 
  mutate_if(is.numeric, round, digits = 1)
shrubs.summary

shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Field Research Station"="Private 1"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Code Farm"="Private 2"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Westminster Ponds"="Public 1"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Five Points Forest"="Private 3"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Medway Valley"="Public 2"))
shrubs.summary$Location <- revalue(shrubs.summary$Location, c("Meadowlily Woods"="Public 3"))

#write.csv(shrubs.summary, file = "shrubs_summary.csv")

```

```{r}
gap.area.summary <- EI5 %>% #summarize gap area
  filter(Category != "Non-Gap") %>%
  group_by(Location, Category) %>% 
  summarize(Gap.Area.Site = mean(Gap.Area) %>% round(1)) 

gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Field Research Station"="Private 1"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Code Farm"="Private 2"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Westminster Ponds"="Public 1"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Five Points Forest"="Private 3"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Medway Valley"="Public 2"))
gap.area.summary$Location <- revalue(gap.area.summary$Location, c("Meadowlily Woods"="Public 3"))

#write.csv(gap.area.summary, file = "gaps_summary.csv")
```



## Run Model Without Western FRS

As per reviewer comments - influence of this station

Remove FRS from data
```{r}
shrubs.EI.wo <- shrubs.EI %>% filter(Location != "Field Research Station") %>% 
  select(Location, Category, ID, Buckthorn, Management, Gap.Area.Site, PC1) 
str(shrubs.EI.wo)
```

Run model without FRS data
```{r}
shrubs.hurdle5.wo <- glmmTMB(Buckthorn ~ Category * PC1 + (1 | Location), 
                        ziformula = ~ Category +  (1 | Location), family = "truncated_nbinom2", data = shrubs.EI.wo)
```

### Compare Model Outputs
```{r}
exp(fixef(shrubs.hurdle5)$cond) # Intercept expected cover when present, others are multiplicative changes in cover conditional on presence
exp(fixef(shrubs.hurdle5.wo)$cond)
```

Remove FRS: 

- Even more buckthorn in average EI EAB gaps
- Greater discrepancy between EAB gaps & non-gaps
- Slope of PC1 in EAB gaps increases

Therefore, greater EAB & EI effect when we remove FRS 


```{r}
#Non-gaps
33.7154190 * 0.2325890
49.2315112 * 0.1975805

#Other gaps
33.7154190 * 0.5355936
49.2315112 *  0.3845417
```

```{r}
exp(fixef(shrubs.hurdle5)$zi) # Intercept is odds of zero, others are odds ratios
exp(fixef(shrubs.hurdle5.wo)$zi) # Intercept is odds of zero, others are odds ratios
```

Remove FRS:

- Still equally likely to be present in EAB & other gaps
- Odds of presense decrease from 1 to 0.76
- Odds of presence in non-gaps also decrease (5.7 to 5.0 times less likely)


```{r}
exp(confint(shrubs.hurdle5))
exp(confint(shrubs.hurdle5.wo))
```

Remove FRS:

- In general, confidence intervals become wider (more uncertainty)

### Visualize Model Predictions

```{r}
pred.wo <- data.frame(PC1 = c(seq(EAB.min, EAB.max, length.out = 100), 
                               seq(other.min, other.max, length.out = 100), 
                                   seq(non.min, non.max, length.out = 100)),
                   Category = c(rep("EAB Gap", 100), rep("Other Gap", 100), rep("Non-Gap", 100)),
                   Gap.Area.Site = rep(mean(shrubs.EI$Gap.Area.Site),300), # Added gap area
                   Management = c(rep("Private", 150), rep("Public", 150)), 
                   Location = rep(NA, 300)) # This NA gives general predictions instead of location specific predictions

pred.wo$fit <- predict(shrubs.hurdle5.wo, newdata = pred, se.fit = TRUE, type = "link")$fit
pred.wo$se <- predict(shrubs.hurdle5.wo, newdata = pred, se.fit = TRUE, type = "link")$se.fit
str(pred.wo)

pred.upr <- pred.wo$fit + 1.96 * pred$se
pred.lwr <- pred.wo$fit - 1.96 * pred$se

pred.wo$upr <- ifelse(exp(pred.upr) < 100, exp(pred.upr), 100)
pred.wo$lwr <- ifelse(exp(pred.lwr) > 0, exp(pred.lwr), 0)

pred.wo$mean <- exp(pred.wo$fit)
```



```{r}
shrubs.EI.wo$Category = factor(shrubs.EI.wo$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))
pred.wo$Category = factor(pred.wo$Category, levels=c("EAB Gap", "Other Gap", "Non-Gap"))

fig.EI.wo <- ggplot(pred.wo, aes(x = PC1, y = mean)) +
    geom_ribbon(aes(x = PC1, ymin = lwr, ymax = upr, fill = Category), alpha = 0.15) +
      geom_point(data = shrubs.EI.wo, aes(x = PC1, y = Buckthorn, col = Category), alpha = 0.6) + 
  geom_line(aes(col = Category), size = 1, alpha = 0.8) +
  labs(x = "Ecological Integrity",
       y = "Buckthorn (%)",
       col = "Category") +
  scale_y_continuous(limits = c(0, 100)) +
    scale_colour_manual(values = mypalette2) +
  scale_fill_manual(values = mypalette2) +
    theme_classic() +
 theme(legend.position = "none")

fig.EI
fig.EI.wo

#ggsave('fig.EI.noFRS.jpeg',fig.EI.wo , units = 'cm', width = 14, height = 10)

```

# Reproducibility
```{r}
Sys.time()
git2r::repository()
sessionInfo()
```

